Realizability-Preserving DG-IMEX Method for the Two-Moment Model of Fermion Transport \tnoteref{support}

### 2018-06-20 Documents/M1/realizableFermionicM1/sections/appendix.tex
Changes due to comments from Tony and Cory.  Work on IMEX section

@@

A second-order accurate, positivity-preserving IMEX scheme of type $A$ (the matrix $A$ is invertible) was given in \cite{hu_etal_2018}.  
We refer to this scheme as IMEX PA2.  
For this scheme, the non-zero components of $\tilde{A}$ and $A$ are given by
\begin{align}
  \tilde{a}_{21} &= 0.7369502715, \nonumber \\
  \tilde{a}_{31} &= 0.3215281691, \quad \tilde{a}_{32} = 0.6784718309, \nonumber \\
  a_{11} &= 0.6286351712, \nonumber \\
  a_{21} &= 0.2431004655, \quad a_{22} = 0.1959392570, \nonumber \\
  a_{31} &= 0.4803651051, \quad a_{32} = 0.0746432814, \quad a_{33} = 0.4449916135. \nonumber
\end{align}
The coefficient in the correction step is $\alpha = 0.2797373792$ and the CFL constant is $c_{sch} = 0.5247457524$.
This scheme is globally stiffly accurate (GSA), so that $\tilde{w}_{i}=\tilde{a}_{3i}$ and $w_{i}=a_{3i}$ for $i\le3$.

@@

We have found another second-order accurate, positivity-preserving IMEX scheme of type $A$, which we refer to as IMEX PA2+.  
This scheme allows for has a larger $c_{sch}$ than IMEX PA2 (i.e., a larger time step while still maintaining positivity).  
The scheme chosen from random sampling of the IMEX parameter space, and was the scheme with the largest $c_{sch}$ that satisfied both the order conditions for second order accuracy and possessed the positivity-preserving property.  
For IMEX PA2+, ($c_{sch} = 0.895041066934$). 
The non-zero components of $\tilde{A}$ and $A$ are given by
\begin{align}
  \tilde{a}_{21} &= 0.909090909090909, \nonumber \\
  \tilde{a}_{31} &= 0.450000000000000, \quad \tilde{a}_{32} = 0.550000000000000, \nonumber \\
  a_{11} &= 0.521932391842510, \nonumber \\
  a_{21} &= 0.479820781424967, \quad a_{22} = 0.002234534340252, \nonumber \\
  a_{31} &= 0.499900000000000, \quad a_{32} = 0.001100000000000, \quad a_{33} = 0.499000000000000. \nonumber
\end{align}
The coefficient in the correction step is $\alpha = 0.260444263529413$.  
This scheme is also GSA; $\tilde{w}_{i}=\tilde{a}_{3i}$ and $w_{i}=a_{3i}$ for $i\le3$.

@@

Another positivity-preserving IMEX scheme was given in \cite{mcclarren_etal_2008} (referred to as a semi-implicit predictor-corrector method in \cite{mcclarren_etal_2008}).  
This scheme can be written in the double Butcher tableau form, and we refer to this scheme as IMEX PC2.  
The non-zero components of $\tilde{A}$ and $A$ are given by
\begin{align}
  \tilde{a}_{21} &= 0.5, \quad \tilde{a}_{32} = 1, \nonumber \\
  a_{22} &= 0.5, \quad a_{33} = 1.0. \nonumber
\end{align}
And $\tilde{w}_{i}=\tilde{a}_{3i} = w_{i}=a_{3i}$ for $i\le3$.

@@

We have found a second-order accurate in the streaming limit, diffusion accurate, positivity-preserving IMEX scheme of type $ARS$, which we refer to as IMEX PARSD.  
This scheme allows for has a $c_{sch}= 1 - 2\epsilon$ with $\epsilon \in [0, 1/2)$.
Here we give an example by setting $\epsilon=0.1$.
\begin{align}
  \tilde{a}_{21} & = 1.0, \nonumber \\
  \tilde{a}_{31} & = 0.5, \quad \tilde{a}_{32} = 0.5, \nonumber \\
  a_{22} & = 1.0, \nonumber \\
  a_{32} & = 0.4 \,( = 0.5 - \epsilon\,), \quad a_{33} = 0.6 \,( = 0.5 + \epsilon\,). \nonumber
\end{align} 
This scheme is also GSA and needs only two compute for implicit terms each time-step.

@@

We compare the performance of the positivity-preserving IMEX schemes with two other, non-positive, schemes.  
The first one is the second-order accurate IMEX scheme given in \cite{cavaglieriBewley2015}.  
We refer to this scheme as IMEX RKCB2.  
The non-zero components of $\tilde{A}$ and $A$ are given by
\begin{align}
  \tilde{a}_{21} &= 2/5, \quad \tilde{a}_{32} = 1, \nonumber \\
  a_{22} &= 2/5, \nonumber \\
  a_{32} &= 5/6, \quad a_{33} = 1/6. \nonumber
\end{align}
And $w_{i} = a_{3i} = \tilde{w}_{i}$ for stiff accuracy.

@@

The second scheme we use for comparison is the second-order accurate IMEX scheme given in \cite{pareschiRusso_2005}.  
We refer to this scheme as IMEX SSP2332.  
The non-zero components of $\tilde{A}$ and $A$ are given by
\begin{align}
  \tilde{a}_{21} &= 1/2, \nonumber \\
  \tilde{a}_{31} &= 1/2, \quad \tilde{a}_{32} = 1/2, \nonumber \\
  a_{11} &= 1/4, \nonumber \\
  a_{22} &= 1/4, \nonumber \\
  a_{31} &= 1/3, \quad a_{32} = 1/3, \quad a_{33} = 1/3. \nonumber
\end{align}
And $w_{i} = a_{3i} = \tilde{w}_{i}$ for stiff accuracy.

@@

\section{Proof of nonexistence of 3-stage Type A ...}

>>>>

For the analysis of positivity-preserving IMEX schemes, additional coefficients are defined \cite{hu_etal_2018} (cf. Eq.~\eqref{eq:imexStagesRewrite}).  
First, let
\begin{equation}
  b_{ii} = \f{1}{a_{ii}}, \quad
  b_{ij} = -\f{1}{a_{ii}}\sum_{l=j}^{i-1}a_{il}b_{lj}, \quad
  \tilde{b}_{ij} = -\f{1}{a_{ii}}\Big(\tilde{a}_{ij}+\sum_{l=j+1}^{i-1}a_{il}\tilde{b}_{lj}\Big).  
\end{equation}
Then, for IMEX schemes of Type~A \cite{dimarcoPareschi2013},
\begin{equation}
  \begin{aligned}
    c_{i0} &= 1-\sum_{j=1}^{i-1}\sum_{l=j}^{i-1}a_{il}b_{lj}, \quad &
    c_{ij} &= \sum_{l=j}^{i-1}a_{il}b_{lj}, \\
    \tilde{c}_{i0} &= 0, \quad &
    \tilde{c}_{ij} &= \tilde{a}_{ij} + \sum_{l=j+1}^{i-1}a_{il}\tilde{b}_{lj}.
  \end{aligned}
\end{equation}
while for IMEX schemes of Type~ARS \cite{ascher_etal_1997}
\begin{equation}
\begin{aligned}
  c_{i0} &= 1-\sum_{j=2}^{i-1}\sum_{l=j}^{i-1}a_{il}b_{lj}, \quad &
  c_{ij} &= \sum_{l=j}^{i-1}a_{il}b_{lj} \\
  \tilde{c}_{i0} &= \tilde{a}_{i1}+\sum_{j=2}^{i-1}a_{ij}\tilde{b}_{j1}, \quad &
  \tilde{c}_{ij} &= \tilde{a}_{ij}+\sum_{l=j+1}^{i-1}a_{il}\tilde{b}_{lj}.  
\end{aligned}
\end{equation}

@@

A second-order accurate, positivity-preserving IMEX scheme of type $A$ (the matrix $A$ is invertible) was given in \cite{hu_etal_2018}.  
We refer to this scheme as IMEX PA2.  
For this scheme, the non-zero components of $\tilde{A}$ and $A$ are given by
\begin{align*}
  \tilde{a}_{21} &= 0.7369502715, \\
  \tilde{a}_{31} &= 0.3215281691, \quad \tilde{a}_{32} = 0.6784718309, \\
  a_{11} &= 0.6286351712, \\
  a_{21} &= 0.2431004655, \quad a_{22} = 0.1959392570, \\
  a_{31} &= 0.4803651051, \quad a_{32} = 0.0746432814, \quad a_{33} = 0.4449916135. 
\end{align*}
The coefficient in the correction step is $\alpha = 0.2797373792$ and the CFL constant is $c_{sch} = 0.5247457524$.
This scheme is globally stiffly accurate (GSA), so that $\tilde{w}_{i}=\tilde{a}_{3i}$ and $w_{i}=a_{3i}$ for $i\le3$.

@@

We have found another second-order accurate, positivity-preserving IMEX scheme of type $A$, which we refer to as IMEX PA2+.  
This scheme allows for has a larger $c_{\mbox{\tiny Sch}}$ than IMEX PA2 (i.e., a larger time step while maintaining positivity).  
The scheme was found by random sampling of the parameter space spanned by the IMEX coefficients, optimizing $c_{\mbox{\tiny Sch}}$.  
For IMEX PA2+, $c_{\mbox{\tiny Sch}} = 0.895041066934$. 
The non-zero components of $\tilde{A}$ and $A$ are given by
\begin{align*}
  \tilde{a}_{21} &= 0.909090909090909, \\
  \tilde{a}_{31} &= 0.450000000000000, \quad \tilde{a}_{32} = 0.550000000000000, \\
  a_{11} &= 0.521932391842510, \\
  a_{21} &= 0.479820781424967, \quad a_{22} = 0.002234534340252, \\
  a_{31} &= 0.499900000000000, \quad a_{32} = 0.001100000000000, \quad a_{33} = 0.499000000000000.
\end{align*}
The coefficient in the correction step is $\alpha = 0.260444263529413$.  
This scheme is also GSA; $\tilde{w}_{i}=\tilde{a}_{3i}$ and $w_{i}=a_{3i}$ for $i\le3$.

The rest of the IMEX schemes we consider here do not include the correction step in Eq.~\eqref{eq:imexCorrection}; i.e., $\alpha=0$.

@@

Another IMEX scheme was given in \cite{mcclarren_etal_2008} (referred to as a semi-implicit predictor-corrector method in \cite{mcclarren_etal_2008}).  
This scheme can be written in the double Butcher tableau form, and we refer to this scheme as IMEX PC2.  
The non-zero components of $\tilde{A}$ and $A$ are given by
\begin{align*}
  \tilde{a}_{21} &= 0.5, \quad \tilde{a}_{32} = 1, \\
  a_{22} &= 0.5, \quad a_{33} = 1.0,
\end{align*}
$\alpha=0$, and $\tilde{w}_{i}=\tilde{a}_{3i} = w_{i}=a_{3i}$ for $i\le3$.

@@

We have found diffusion accurate, positivity-preserving IMEX schemes of type ARS that are second-order accurate in the streaming limit, which we refer to as IMEX PARSD; see \ref{app:PARSD}.  
For these schemes, $c_{\mbox{\tiny Sch}}= 1 - 2\epsilon$ with $\epsilon \in [0, 1/2)$.
Here we give an example by setting $\epsilon=0.1$.
\begin{align*}
  \tilde{a}_{21} & = 1.0, \\
  \tilde{a}_{31} & = 0.5, \quad \tilde{a}_{32} = 0.5, \\
  a_{22} & = 1.0, \nonumber \\
  a_{32} & = 0.4 \,( = 0.5 - \epsilon\,), \quad a_{33} = 0.6 \,( = 0.5 + \epsilon\,). 
\end{align*}
This scheme is GSA, $\alpha=0$, and requires only two implicit solves per time step.

@@

We compare the performance of the positivity-preserving IMEX schemes with two other, non-positive, schemes.  
The first one is the second-order accurate IMEX scheme given in \cite{cavaglieriBewley2015}.  
We refer to this scheme as IMEX RKCB2.  
The non-zero components of $\tilde{A}$ and $A$ are given by
\begin{align*}
  \tilde{a}_{21} &= 2/5, \quad \tilde{a}_{32} = 1, \\
  a_{22} &= 2/5, \nonumber \\
  a_{32} &= 5/6, \quad a_{33} = 1/6,
\end{align*}
$\alpha=0$, and $w_{i} = a_{3i} = \tilde{w}_{i}$ (stiffly accurate \cite{pareschiRusso_2005}).

@@

Another scheme we use for comparison is the second-order accurate IMEX scheme given in \cite{pareschiRusso_2005}.  
We refer to this scheme as IMEX SSP2332.  
The non-zero components of $\tilde{A}$ and $A$ are given by
\begin{align*}
  \tilde{a}_{21} &= 1/2, \\
  \tilde{a}_{31} &= 1/2, \quad \tilde{a}_{32} = 1/2, \\
  a_{11} &= 1/4, \\
  a_{22} &= 1/4, \\
  a_{31} &= 1/3, \quad a_{32} = 1/3, \quad a_{33} = 1/3, 
\end{align*}
$\alpha=0$, and $w_{i} = a_{3i} = \tilde{w}_{i}$ (stiffly accurate).

@@

\section{Proof of nonexistence of 3-stage Type A ...}
\label{app:noTypeA}

### 2018-06-20 Documents/M1/realizableFermionicM1/sections/imex.tex
Changes due to comments from Tony and Cory.  Work on IMEX section

(The realizability-preserving and positivity-preserving properties of an IMEX scheme are intimately related, and we use two phrases interchangeably.)  

@@

To prove the positivity-preserving property of the IMEX scheme, Hu et al. \cite{hu_etal_2018} rewrite the stage values in Eq.~\eqref{imexStages} in the following form
\begin{equation}
  \vect{u}^{(i)}
  =\sum_{j=0}^{i-1}c_{ij}\Big[\,\vect{u}^{(j)}+\hat{c}_{ij}\,\dt\,\vect{\cT}(\vect{u}^{(j)})\,\Big]
  +a_{ii}\,\dt\,\f{1}{\tau}\,\vect{\cQ}(\vect{u}^{(i)}),\quad i=1,\ldots,s,
  \label{eq:imexStagesRewrite}
\end{equation}
where $c_{ij}$, and $\hat{c}_{ij}=\tilde{c}_{ij}/c_{ij}$ are computed from $\tilde{a}_{ij}$ and $a_{ij}$.
Such that, $c_{i0} = 1 - \sum_{j=1}^{i-1}c_{ij}$, $c_{ij} = \sum_{l=j}^{i-1}a_{il}b_{lj}$, $\tilde{c}_{i0} = \tilde{a}_{i1} + \sum_{j=2}^{i-1} a_{ij}\tilde{b}_{j1}$,$\tilde{c}_{ij} = \tilde{a}_{ij} + \sum_{l=j+1}^{i-1} a_{il}\tilde{b}_{lj}$ $b_{ij} = - \frac{1}{a_{ii}} \sum_{l = j}^{i-1}a_{il}b_{lj}$, $b_{ii} = \frac{1}{a_{ii}}$, $\tilde{b}_{ij} = \frac{1}{a_{ii}}\big( -\tilde{a}_{ij} - \sum_{l = j+1}^{i-1}a_{il}\tilde{b}_{lj}\big)$. 
(In Eq.~\eqref{eq:imexStagesRewrite}, $\vect{u}^{(0)}=\vect{u}^{n}$.)  
Two types of IMEX schemes are considered: \emph{type~A} \cite{pareschiRusso_2005} and \emph{type~ARS} \cite{ascher_etal_1997}.  
For IMEX schemes of type~A, the matrix $A$ is invertible.  
For IMEX schemes of type~ARS, the matrix $A$ can be written as
\[\left( 
   \begin{matrix} 
       0 & 0 \\ 
       0 & \hat{A}
   \end{matrix}
\right)\]
where $\hat{A}$ is invertible.
In writing the stages in the IMEX scheme in the general form given by Eq.~\eqref{eq:imexStagesRewrite}, it should be noted that $\tilde{c}_{i0}=0$ for IMEX schemes of Type~A, and $c_{i1}=\tilde{c}_{i1}=0$ for IMEX schemes of Type~ARS \cite{hu_etal_2018}.  
For type A schemes, the positivity-preserving property follows from requiring $a_{ii}>0$ and $c_{i0}\ge0$ for $i=1,\ldots,s$, and $c_{ij},\tilde{c}_{ij}\ge0$, for $i=2,\ldots,s$, $j=1,\ldots,s-1$.  
For type ARS schemes, the positivity follows from requiring $a_{ii}>0$ and $c_{i0},\tilde{c}_{i0}\ge0$ for $i=2,\ldots,s$, and $c_{ij},\tilde{c}_{ij}\ge0$ for $i=3,\ldots,s$, $j=2,\ldots,i-1$.  
Such coefficients were given in \cite{hu_etal_2018} for GSA schemes of type A with $s=3$ and type ARS with $s=4$.  
(It was also proven that $s=3$ and $s=4$ are the necessary number of stages needed for GSA second-order positivity-preserving IMEX schemes of type A and type ARS, respectively.)

Importantly, the coefficients $c_{ij}$ in Eq.~\eqref{eq:imexStagesRewrite} satisfy $\sum_{j=0}^{i-1}c_{ij}=1$, which implies the so-called Shu-Osher form.  
Thus, if the expression inside the square brackets in Eq.~\eqref{eq:imexStagesRewrite} --- which is in the form of a forward Euler update with time step $\hat{c}_{ij}\,\dt$ --- is positive for all $i=1,\ldots,s$, $j=0,\ldots,i-1$, positivity of the entire sum on the right-hand side of Eq.~\eqref{eq:imexStagesRewrite} follows from convexity arguments.  
Thus, if the explicit update with the transport operator is positive for a time step $\dt_{\mbox{\tiny Ex}}$, the IMEX scheme is positivity-preserving for a time step $\dt\le c_{\mbox{\tiny Sch}}\,\dt_{\mbox{\tiny Ex}}$, where
\begin{equation}
  c_{\mbox{\tiny Sch}}=\min_{\substack{i = 2,\ldots,s \\ 
       j = 1,\ldots,i-1}}\,\f{1}{\hat{c}_{ij}} \quad \text{(Type A)}, \quad  c_{\mbox{\tiny Sch}}=\min_{\substack{i = 2,\ldots,s \\ 
              j = 0,2,\ldots,i-1}}\,\f{1}{\hat{c}_{ij}} \quad \text{(Type ARS)}.
  \label{eq:imexCFL}
\end{equation}
It is desirable to make $c_{\mbox{\tiny Sch}}$ as large (close to $1$) as possible.  
(Positivity of $\vect{u}^{(i)}$ also requires the implicit solve to preserve positivity.)  
In \cite{hu_etal_2018}, Hu et al. provide examples of GSA, positivity-preserving IMEX schemes of type A (see scheme PA2 in \ref{app:butcherTables}) and type ARS.  
In \ref{app:butcherTables}, we provide another example of a GSA, positivity-preserving IMEX scheme of type A (scheme PA2+), with a significantly larger $c_{\mbox{\tiny Sch}}$ (a factor of about $1.7$ larger).

@@

Unfortunately, the correction step in Eq.~\eqref{eq:imexCorrection} deteriorates the accuracy of the IMEX scheme when applied to the moment equations in the diffusion limit ($\xi=0$, $\tau\ll 1$), where $\vect{\cK}\approx\f{1}{3}\,\cJ\,\vect{I}$.  
In the diffusion limit, to leading order in $\tau$, we then have $\vect{\cH}\approx-\f{1}{3}\,\tau\,\nabla\cJ$, which is due to a balance between the transport term and the collision term in the equation for the particle flux $\vect{\cH}$.  
The absence of the transport operator in the correction step, destroys this balance.  
We demonstrate the inferior performance of IMEX schemes with the correction step given by Eq.~\eqref{eq:imexCorrection} in the diffusion limit in Section~\ref{sec:smoothProblems}.  
We have also implemented and tested one of the IMEX schemes in Chertock et al. \cite{chertock_etal_2015}, where the transport operator is included in the correction step, and found it to perform very well in the diffusion limit.  
However, we have not been able to prove sufficiently general realizability-preserving properties with this approach, without invoking a too severe time step restriction.  
We therefore proceed to design positivity-preserving IMEX schemes, without the correction step, that perform better in the diffusion limit.  
We limit the scope to IMEX schemes of type~ARS.  
(It can be shown that IMEX schemes of type~A satisfying the conditions we list below Eq.~\eqref{eq:orderConditionsEx} do not exist.)

@@

  \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}=\f{1}{2}.  

@@

We then seek to design GSA IMEX schemes of Type~ARS with the following properties

@@

@@

Here we analyze the absolute stability of the proposed IMEX schemes, PA2+ and PARSD (given in \ref{app:butcherTables}), following \cite{hu_etal_2018}.  
As is commonly done, we do this in the context of the linear scalar equation
\begin{equation}
  \dot{u}=\lambda_{1}\,u+\lambda_{2}\,u,
  \label{eq:scalarODE}
\end{equation}
where $\lambda_{1}\in\mathbb{C}$ and real $\lambda_{2}\le0$.  
On the right-hand side of Eq.~\eqref{eq:scalarODE}, the first (oscillatory) term is treated explicitly, while the second (damping) term is treated implicitly.  
The IMEX schemes can then be written as $u^{n+1} =P(z_{1},z_{2})\,u^{n}$, where $P(z_{1},z_{2})$ is the amplification factor of the scheme, $z_{1}=\dt\,\lambda_{1} = x +iy$, and $z_{2} = \dt\,\lambda_{2} < 0$.  
Stability of the IMEX scheme requires $|P(z_1,z_2)|\leq 1$.  
The stability regions of PA2+ and PARSD are plotted in Figure~\ref{fig:AbsoluteStability}.  
As can be seen from the figures, the absolute stability region of both schemes increases with increasing $|z_{2}|$.  
For a given $|z_{2}|>0$, the stability region of PARSD is larger than that for PA2+.  
In the linear case, a time step that satisfies the absolute stability for the explicit part of the IMEX scheme ($|z_2| = 0$) fulfills the stability requirement for the IMEX scheme as whole ($|z_2| \geq 0$).  
This conclusion is valid for all PARSD schemes with $\epsilon\in [0,0.5)$.
\begin{figure}[h]
  \centering
  \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{figures/AbsoluteStabilityPARSD}
    \includegraphics[width=0.5\textwidth]{figures/AbsoluteStabilityPA2+}
  \end{tabular}
   \caption{Boundary of the absolute stability region for different values of $z_2$ for PARSD with $\epsilon = 0.1$ (left panel) and PA2+ (right panel).}
  \label{fig:AbsoluteStability}
\end{figure}

>>>>

(The realizability-preserving and positivity-preserving properties of an IMEX scheme are intimately related, and we use the two phrases interchangeably.)  

@@

To prove the positivity-preserving property of the IMEX scheme, Hu et al. \cite{hu_etal_2018} rewrite the stage values in Eq.~\eqref{imexStages} in the following form
\begin{equation}
  \vect{u}^{(i)}
  =\sum_{j=0}^{i-1}c_{ij}\Big[\,\vect{u}^{(j)}+\hat{c}_{ij}\,\dt\,\vect{\cT}(\vect{u}^{(j)})\,\Big]
  +a_{ii}\,\dt\,\f{1}{\tau}\,\vect{\cQ}(\vect{u}^{(i)}),\quad i=1,\ldots,s,
  \label{eq:imexStagesRewrite}
\end{equation}
where $c_{ij}$, and $\hat{c}_{ij}=\tilde{c}_{ij}/c_{ij}$ are computed from $\tilde{a}_{ij}$ and $a_{ij}$ (see \ref{app:butcherTables}).  
In Eq.~\eqref{eq:imexStagesRewrite}, $\vect{u}^{(0)}=\vect{u}^{n}$.  
Two types of IMEX schemes are considered: \emph{type~A} \cite{pareschiRusso_2005,dimarcoPareschi2013} and \emph{type~ARS} \cite{ascher_etal_1997}.  
For IMEX schemes of type~A, the matrix $A$ is invertible.  
For IMEX schemes of type~ARS, the matrix $A$ can be written as
\[\left( 
   \begin{matrix} 
       0 & 0 \\ 
       0 & \hat{A}
   \end{matrix}
\right)\]
where $\hat{A}$ is invertible.  
In writing the stages in the IMEX scheme in the general form given by Eq.~\eqref{eq:imexStagesRewrite}, it should be noted that $\tilde{c}_{i0}=0$ for IMEX schemes of Type~A, and $c_{i1}=\tilde{c}_{i1}=0$ for IMEX schemes of Type~ARS \cite{hu_etal_2018}.  
For type A schemes, the positivity-preserving property follows from requiring $a_{ii}>0$ and $c_{i0}\ge0$ for $i=1,\ldots,s$, and $c_{ij},\tilde{c}_{ij}\ge0$, for $i=2,\ldots,s$, $j=1,\ldots,s-1$.  
For type ARS schemes, positivity follows from requiring $a_{ii}>0$ and $c_{i0},\tilde{c}_{i0}\ge0$ for $i=2,\ldots,s$, and $c_{ij},\tilde{c}_{ij}\ge0$ for $i=3,\ldots,s$, $j=2,\ldots,i-1$.  
Such coefficients were given in \cite{hu_etal_2018} for GSA schemes of type A with $s=3$ and type ARS with $s=4$.  
(It was also proven that $s=3$ and $s=4$ are the necessary number of stages needed for GSA second-order positivity-preserving IMEX schemes of type A and type ARS, respectively.)

Importantly, the coefficients $c_{ij}$ in Eq.~\eqref{eq:imexStagesRewrite} satisfy $\sum_{j=0}^{i-1}c_{ij}=1$, which implies the so-called Shu-Osher form of the explicit part \cite{shuOsher_1988}.  
If the expression inside the square brackets in Eq.~\eqref{eq:imexStagesRewrite} --- which is in the form of a forward Euler update with time step $\hat{c}_{ij}\,\dt$ --- is positive for all $i=1,\ldots,s$, $j=0,\ldots,i-1$, positivity of the entire sum on the right-hand side of Eq.~\eqref{eq:imexStagesRewrite} follows from convexity arguments.  
Thus, if the explicit update with the transport operator is positive for a time step $\dt_{\mbox{\tiny Ex}}$, the IMEX scheme is positivity-preserving for a time step $\dt\le c_{\mbox{\tiny Sch}}\,\dt_{\mbox{\tiny Ex}}$, where
\begin{equation}
  c_{\mbox{\tiny Sch}}=\min_{\substack{i = 2,\ldots,s \\ 
       j = 1,\ldots,i-1}}\,\f{1}{\hat{c}_{ij}} \quad \text{(Type A)}, \quad  c_{\mbox{\tiny Sch}}=\min_{\substack{i = 2,\ldots,s \\ 
              j = 0,2,\ldots,i-1}}\,\f{1}{\hat{c}_{ij}} \quad \text{(Type ARS)}.
  \label{eq:imexCFL}
\end{equation}
It is desirable to make $c_{\mbox{\tiny Sch}}$ as large (close to $1$) as possible.  
Note that positivity of $\vect{u}^{(i)}$ also requires the implicit solve (equivalent to implicit Euler) to preserve positivity.  
In \cite{hu_etal_2018}, Hu et al. provide examples of GSA, positivity-preserving IMEX schemes of type A (see scheme PA2 in \ref{app:butcherTables}) and type ARS.  
In \ref{app:butcherTables}, we provide another example of a GSA, positivity-preserving IMEX scheme of type A (scheme PA2+), with a larger $c_{\mbox{\tiny Sch}}$ (a factor of about $1.7$ larger).

@@

Unfortunately, the correction step in Eq.~\eqref{eq:imexCorrection} deteriorates the accuracy of the IMEX scheme when applied to the moment equations in the diffusion limit ($\xi=0$, $\tau\ll 1$), where $\vect{\cK}\approx\f{1}{3}\,\cJ\,\vect{I}$.  
In the diffusion limit, to leading order in $\tau$, we then have $\vect{\cH}\approx-\f{1}{3}\,\tau\,\nabla\cJ$, which is due to a balance between the transport term and the collision term in the equation for the particle flux $\vect{\cH}$.  
The absence of the transport operator in the correction step, destroys this balance.  
We demonstrate the inferior performance of IMEX schemes with the correction step given by Eq.~\eqref{eq:imexCorrection} in the diffusion limit in Section~\ref{sec:smoothProblems}.  
We have also implemented and tested one of the IMEX schemes in Chertock et al. \cite{chertock_etal_2015}, where the transport operator is included in the correction step, and found it to perform very well in the diffusion limit.  
However, we have not been able to prove the realizability-preserving property with this approach without invoking a too severe time step restriction.  
We therefore proceed to design positivity-preserving IMEX schemes, without the correction step, that perform better in the diffusion limit.  
We limit the scope to IMEX schemes of type~ARS.  
(It can be shown that IMEX schemes of type~A conforming to Definition~\ref{def:DIMEXP} below do not exist; cf. \ref{app:noTypeA}.)

@@

  \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}=\f{1}{2},

@@

where the first condition (consistency condition) is required for first-order accuracy.  
We then seek to design GSA IMEX schemes of Type~ARS conforming to the following definition
\begin{define}
  A diffusion accurate, positivity-preserving IMEX scheme satisfies the following properties

@@

  \label{def:DIMEXP}
\end{define}

@@

Here we analyze the absolute stability of the proposed IMEX schemes, PA2+ and PARSD (given in \ref{app:butcherTables}), following \cite{hu_etal_2018}.  
As is commonly done, we do this in the context of the linear scalar equation
\begin{equation}
  \dot{u}=\lambda_{1}\,u+\lambda_{2}\,u,
  \label{eq:scalarODE}
\end{equation}
where $\lambda_{1}\in\mathbb{C}$ and real $\lambda_{2}\le0$.  
On the right-hand side of Eq.~\eqref{eq:scalarODE}, the first (oscillatory) term is treated explicitly, while the second (damping) term is treated implicitly.  
The IMEX schemes can then be written as $u^{n+1} =P(z_{1},z_{2})\,u^{n}$, where $P(z_{1},z_{2})$ is the amplification factor of the scheme, $z_{1}=\dt\,\lambda_{1} = x +iy$, and $z_{2} = \dt\,\lambda_{2} < 0$.  
Stability of the IMEX scheme requires $|P(z_1,z_2)|\leq 1$.  
The stability regions of PA2+ and PARSD are plotted in Figure~\ref{fig:AbsoluteStability}.  
As can be seen from the figures, the absolute stability region of both schemes increases with increasing $|z_{2}|$.  
For a given $|z_{2}|>0$, the stability region of PARSD is larger than that of PA2+.  
In the linear case, a time step that satisfies the absolute stability for the explicit part of the IMEX scheme ($|z_2| = 0$) fulfills the stability requirement for the IMEX scheme as whole ($|z_2| \geq 0$).  
This conclusion is valid for all PARSD schemes with $\epsilon\in [0,0.5)$.
\begin{figure}[h]
  \centering
  \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{figures/AbsoluteStabilityPARSD}
    \includegraphics[width=0.5\textwidth]{figures/AbsoluteStabilityPA2+}
  \end{tabular}
   \caption{Boundary of the absolute stability region for different values of $z_2$ for PARSD with $\epsilon = 0.1$ (left panel) and PA2+ (right panel).}
  \label{fig:AbsoluteStability}
\end{figure}
